---
title: "Simple R Functions "
author: "Ziran Min"
date: "January 26, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

******

Here is \texttt{tmpFn1}
```{r}

tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}

## simple example
a <- c(2, 5, 3, 8, 2, 4)

b <- tmpFn1(a)
b

```

and now \texttt{tmpFn2}

```{r}
tmpFn2 <- function(xVec2){
  
  n = length(xVec2)
  
  return(xVec2^(1:n)/(1:n))
}

c <- tmpFn2(a)
c
```

######  
  
  
  
(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$
```{r}
tmpFn3 <- function(x, n){
  return (1 + sum((x^(1:n))/(1:n)))
}
d <- tmpFn3(2, 5)
d
```
******

\newpage
### 2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$

Try out your function.  \texttt{tmpFn}(c(1:5,6:1))
******
```{r}
tmpFn <- function(xVec){
  n <- length(xVec)
  first <- (1:(n-2))
  second <- first + 1
  third <- first + 2
  return ((xVec[first] + xVec[second] + xVec[third])/3)
}
e <- tmpFn(c(1:5, 6:1))
e
```
******
### 3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}.  
Hence plot the function $f(x)$ for $-3 < x <3$.
******
```{r}
tmpFn <- function(x){
  ifelse( (x < 0), (x^2 + 2*x + 3),
          ifelse( x < 2, x + 3, x^2 + 4*x - 7)
  )
}
tmp <- seq(-3, 3, len=10000)
plot(tmp, tmpFn(tmp), type = "l", xlab = "x", ylab = "f(x)")
```
******
### 4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  
Hence the result of using the function on the matrix   

$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

should be:

$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$
******
```{r}
tmpFn <- function(m){
  m[m%%2 == 1] <- 2* m[m%%2 == 1]
  m
}

test <- matrix(c(1,1,3,5,2,6,-2,-1,-3), nrow=3, byrow=TRUE)
test
tmpFn(test)
```
******
### 5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:


$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$
******
```{r}
tmpFn <- function(n, k) {
  tmp <- diag(k, nrow = n)
  tmp[abs(row(tmp) - col(tmp)) == 1] <- 1
  tmp
}
tmpFn(6,3)
```
******
### 6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   

Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.
******
```{r}
quadrant <- function(alpha) {
 1 + (alpha %% 360) %/% 90
}
quadrant(45)
quadrant(105)
quadrant(250)
quadrant(300)
quadrant(1000)
```
******
### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

$k$ = the day of the month  
$y$ = the year in the century  
$c$ = the first 2 digits of the year (the century number)  
$m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   
For example, the date 21/07/1`963 has $m = 5, k = 21, c = 19, y = 63$;  
the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  
Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year.  
Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.
******
```{r}
weekday <- function(day, month, year) {
 k <- day
 m <- month - 2
  if (m <= 0) {
    m <- m + 12
    year <- year - 1
  }

 c <- year %/% 100 
 y <- year %% 100
 tmp <- (floor(2.6*m - 0.2) + k + y + y %/% 4 + c %/% 4 - 2 * c) 
 week <- c("Sunday","Monday","Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
 return (week[tmp%% 7 + 1])
}
weekday(30,1,2018)
```
******
(b) Does your function work if the input parameters \text{day, month, and year} are vectors with the same length and valid entries?
******
```{r}
days <- c(1,6,10)
months <- c(2,2,2)
years <- c(2018,2018,2018)
weekday(days, months,years)
# This returns "the condition has length > 1 and only the first element will be used" but still gives me anwers, so my function works
```
******

### 8.

(a) Suppose $x_0 = 1$ and $x_1 = 2$ and $$x_j = x_{j-1} + 2 / x_{j-1}$$ for $j = 1,2,...$

Write a function testLoop(n) that returns the first n - 1 elements of the sequence
*****
```{r}
testLoop <- function(n){ 
  x <- rep(0, n-1)
  x[1] <- 1
  x[2] <- 2
  for(i in 3:length(x)){
    x[i] <- x[i-1] + 2/x[i-1]
  }
  return(x)
}
#test 
testLoop(10)
```

(b) define testLoop2 that calculates $$\sum_{j=1}^{n}e^j$$ for a vector y

```{r}
testLoop2 <- function(yVec){ 
  n <- length(yVec)
  tmp <- exp(1:n)
  return(sum(tmp))
}
testLoop2(c(1:8)) 
testLoop2(c(2,59,8,3,10,34,55,4)) 
```
*****

###9.
(a)
*****
```{r}
quadmap <- function(start, rho, niter){
  x <- rep(0, niter)
  x[1] <- start
  for(i in 2:niter){
    x[i] = rho * x[i-1] * (1- x[i-1])
  }
  return(x)
}

quadmap(start=0.8, rho=2, niter=100) 
tmp<-quadmap(start=0.95, rho=2.99, niter=500)
plot(tmp, type="l")
plot(tmp[300:500], type="l")
```

(b)
```{r}
quadmap2 <- function(start, rho){ 
  x1 <- start
  n <- 1
  while(abs(x1 - rho * x1 * (1 - x1)) >= 0.02 ){ 
    x1 <- rho * x1 * (1 - x1)
    n <- n + 1
  }
  return(n)
}
quadmap2(start=.95, rho=2.99) 
```
*****

###10. (a) Given a vector $(x_1, ... , x_n)$ the sample autocorrelation of lag k is defined to be $$r_k = \frac{\sum_{i=k+1}^n(x_i - x_s)(x_{i-k} - x_s)}{\sum_{i=k+1}^n(x_i - x_s)^2}$$ Where $x_s$ is the sample mean write a function to solve for r1 and r2
*****
```{r}
tmpFn <- function(xVec){
  x <- xVec - mean(xVec)
  sxx <- sum(x^2)
  n <- length(xVec)
  r1 <- sum ( x[2:n] * x[1:(n-1)] ) / sxx
  r2 <- sum ( x[3:n] * x[1:(n-2)] ) / sxx
  return (c(r1, r2))
}

#testing function over given interval
tmpFn(seq(2,56, by=3))
```

since I started out with nearly the generalised code, It is simple to allow for my result to list $r_0, ..., r_k$ (b)

```{r}
tmpFn2 <- function(xVec, k){
  x <- xVec - mean(xVec)
  sxx <- sum(x^2)
  n <- length(x)
  return(c(1, sapply((1:k), function(a) sum( x[(a+1):n] * x[1:(n-a)] / sxx  ))))
}

tmpFn2(seq(2,56, by=3), 5)
```
*****









